### 2023-06-05 manuscript.Rmd
revision based on peer review comments

---
title             : "Commentary: 'Responsible Research Assessment II: A specific proposal for hiring and promotion in psychology'"
shorttitle        : "Responsible research assessment"

@@

At present, whether a piece of software is open source is not evaluated in the pre-screening phase.
However, an open license is central for assessing both rigor and (potential) impact, and should be part of the phase I assessment.
Above, we already discussed the different ways research software can have an impact --- not only by direct usage, but also by reusing software in other packages.
Open-source software makes broader impact more likely and is a prerequisite for full transparency and reproducibility [@peikert2021reproducible].
In addition, many aspects of rigor are impossible to evaluate for closed-source software --- for example, whether it is well-tested or bugs have been fixed.
Therefore, we propose to penalize software if it does not adhere to an open-source license (those approved in a review process by the Open Source Initiative (https://opensource.org/licenses) by allowing them only half of the total achievable points.

@@

We believe that the suggested merit statement is most useful to assess impact of research software and that this should be the primary statement for committee members to evaluate if they are less concerned with the technical aspects of research software development.
In our view, researchers should be requested to indicate at least one (and up to three) research projects that directly benefited from their software contributions.
We believe this to be a fair assessment of the actual impact the specific contribution of the individual researcher had.

@@

In sum, we are thankful to @gartner2022responsible for highlighting the importance of research software contributions as scientifically valuable products.
We believe the current proposal should aim to better reflect the distinction between rigor and impact for software, similar to the guidelines proposed to evaluate journal articles.
To this end, we suggest a more fine-grained assessment of rigor, and put emphasis on the merit statement, in which software authors should argue in how far their project impacted other scientific endeavors.
Last, we hope to arrive at an evaluation scheme that incentivizes the development of scientific open-source software.

>>>>

---
title             : "Assessing rigor and impact in research software"
#subtitle          : "Commentary: 'Responsible Research Assessment II: A specific proposal for hiring and promotion in psychology'"
shorttitle        : "Rigor and impact"

@@

At present, whether a piece of software is open source is not evaluated at the pre-screening stage.
However, an open license is central for assessing both rigor and (potential) impact, and should be part of the phase I assessment.
Above, we already discussed the different ways research software can have an impact --- not only by direct usage, but also by reusing software in other packages.
Open-source software makes broader impact more likely and is a prerequisite for full transparency and reproducibility [@peikert2021reproducible].
In addition, many aspects of rigor are impossible to evaluate for closed-source software --- for example, whether it is well-tested or bugs have been fixed.
Giving more weight to open licenses aligns well with the original article's emphasis on open science.
Therefore, we propose to penalize software if it does not adhere to an open-source license (those approved in a review process by the Open Source Initiative (https://opensource.org/licenses), for example, by allowing them only half of the total achievable points.

@@

We believe that the suggested merit statement is most useful to assess impact of research software and that this should be the primary statement for committee members to evaluate if they are less concerned with the technical aspects of research software development.
In our view, researchers should be requested to indicate at least one (and up to three) research projects that directly benefited from their software contributions.
In case of recently published research software, the merit statement can be used alternatively to describe expected benefits to the field or potential use-cases of the software.
We believe this to be a fair assessment of the actual impact the specific contribution of the individual researcher had.

@@

In sum, we are thankful to @gartner2022responsible for highlighting the importance of research software contributions as scientifically valuable products.
We believe the current proposal should aim to better reflect the distinction between rigor and impact for software, similar to the guidelines proposed to evaluate journal articles.
To this end, we suggest a more fine-grained assessment of rigor, and put emphasis on the merit statement, in which software authors should argue in how far their project impacted other scientific endeavors.
Last, we hope to arrive at an evaluation scheme that incentivizes the development of scientific open-source software.
Note that the requirements of a given academic position should guide the weighting of research software when evaluating and selecting candidates. In some cases, research software development may be central to a position and therefore a requirement (such as when hiring professors of statistics or research methods), but in other cases, we hope that research software development activities at least are considered as research output comparable to published papers.
